Standard Candle Cosmology With Incomplete Spectroscopic Classification

### 2015-09-28 doc/article.tex
adding to paper and reformatting of table

\documentclass[preprint]{elsarticle}
\usepackage{aas_macros}
\usepackage{amsmath,amssymb}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{hyperref}

@@

Photometric SN analysis has been developed
in the hope of taking advantage of the statistical weight imported by the
availability of light curves
for large numbers of supernovae, despite their lack of spectroscopy.
Such analyses have been applied to SDSS-II SN data \citep{2012ApJ...752...79H,
2013ApJ...763...88C}, and are planned for
DES \citep{2012ApJ...753..152B} and LSST \citep{2012arXiv1211.0310L}.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=7in]{/Users/akim/project/abc/results/hdpgm.pdf} 
   \caption{Probabilistic Graphical Model for the SN~Ia analysis.  
   The forward modeling
   flow goes from left to right, starting with the description of the universe, observatory,
   and data.    A transient has host galaxy $g_i$, which determines its redshift $z_i$
   and galaxy parameters $\theta_{gi}$.
   A model with parameters $\theta_\mu$ fix the distance modulus $\mu_i$.
   The transient type $T$ depends on the host-galaxy parameters  $\theta_{gi}$
   and rates $\theta_r$.   The transient
   parameters $\theta_T^X$, $\theta_{Ti}^X$ (and perhaps the host galaxy) determine the luminosity $L$.       The 
   incoming photon flux $n$, $n_g$  are then fixed
   with redshift and distance modulus.
   The instrumental transmission function $\phi$ is calibrated with data ${Z}$ and
   gives the expected
   counts $\overline{\mathit{ADU}}$, $\overline{\mathit{ADU}}_g$. 
   The transient has realized light curves (${ADU}$) and passes selection criteria
   through $\tau_i$ to be included in the sample.  Some transients have spectral data
   ${X}_S$.  The properties of the galaxy identified as the host galaxy is $\theta_G$. 
   \label{pgm:fig}}
\end{figure}

From these fundamental parameters we can determine the expected
signals
through a series of intermediary parameters,
some of which are deterministic (points in Figure~\ref{pgm:fig}) and others
probabilistic (ovals).
\begin{itemize}
\item $z_i$, $\theta_{gi}$ are the galaxy redshift and parameters (e.g.\ mass, sSFR, metallicity).
Each galaxy $g_i$ has its values for these quantities.
\item $\mu_i=\mu(z_i; \theta_\mu)$.  The distance modulus is a function of redshift,
Unspecified at the moment,  it could
be a $\Lambda$CDM prediction with cosmological parameters, a spline with knot values
as parameters.
The redshift of the galaxy is used in the Hubble diagram, peculiar velocities are not considered.
\item $P(T_i | \theta_r, \theta_{gi})$.  The transient type $T_i$, SN~Ia or non-Ia, depends
on the relative rates given the host properties.
\item $P(L_i(t,\lambda)| T_i, \theta_T^{Ia}, \theta_{Ti}^{Ia}, \theta_T^{\mathit{non-Ia}}, \theta_{Ti}^{\mathit{non-Ia}},
\theta_{gi})$.  The source model determines
the luminosity; the model may include intrinsic dispersion to make the luminosity
probabilistic. The  model includes  information on the
source and line-of-sight effects, which specify the effective SED.   Only the
parameters of the model of the appropriate type $T$ are relevant.  The galaxy parameters
$\theta_{gi}$ are included if correlations with host-galaxy properties are to be modeled.
\item $n_i(t,\lambda; L_i, \mu_i, z_i)$, $n_{gi}(t; g_i, \mu_i, z_i)$.  The  fluxes of
the transient and host that are incident at Earth
are fixed by the luminosity, distance modulus, and redshift.
The supernova redshift is required for this calculation,
for simplicity the supernova and host-galaxy redshift are assumed to be equal.
\item $\overline{\mathit{ADU}}$ and
$\overline{\mathit{ADU}}_g$ are the expected counts of the transient and galaxy respectively,
coming from $\int n_i \phi d\lambda$.  
\end{itemize}

The observations are as follows:
\begin{itemize}
\item ${\mathit{ADU}}$ is the realized flux.  It depends on the underlying flux and
the photometric noise.
\item $\theta_G$ represents the galaxy parameters of the
galaxy catalog used to identify prospective host galaxies.
\item ${T}_S$, ${z}_S$, ${\theta}_S$ are the transient type, redshift and properties from
spectroscopic data. Details of the actual measurement are ignored and a direct association
is made with the fluxes of the transient and underlying galaxy $n$, $n_g$.
\item $\tau$ is the score used to decide if the transient is used in the analysis.  It is
determined based on other measurements $\tau(\mathit{ADU},T_S)$.
\item ${Z}$ represents the calibration measurements.   Although this may be given
by the Calibration Task Force, the PGM is written for the more general case where the calibration
is modeled within our analysis.
\end{itemize}

\section{Simple Example}
\subsection{Model}
\label{example:sec}
A simplified version of the above model,
\begin{equation}
p(\mathit{ADU}, {{T}}_S,{{z}}_S, \theta_G|  \Omega_M, w, \theta_r,\alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}}),
\end{equation}
is examined in this section and is sketched in Figure~\ref{toypgm:fig}.
The measurements are
\begin{itemize}
\item $\mathit{ADU}$: the measured counts at peak brightness;
\item ${{T}}_S$, ${{z}}_S$: the spectroscopic type and redshift for a subset of transients;
\item $\theta_{G1}$, $\theta_{G2}$: the redshifts of the identified host and neighbor
galaxies.
\end{itemize}
The model parameters are
\begin{itemize}
\item $\Omega_M$, $w$: the parameters of a flat dark-energy cosmology;
\item $\theta_{r1}$, $\theta_{r2}$:  for two redshifts the fraction of transients that are SN~Ia;
\item $\alpha_{Ia}$, $\sigma_{Ia}$, $\alpha_{\mathit{non-Ia}}$, $\sigma_{\mathit{non-Ia}}$:
the luminosity (times $4\pi$) and intrinsic dispersion for SNe~Ia and non-Ia at peak.
Light-curve shape and color are not considered in this simple model.
\end{itemize}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5in]{/Users/akim/project/abc/results/toy_pgm.pdf} 
   \caption{Probabilistic Graphical Model for the simple SN~Ia analysis. The
   parameters include the cosmological parameters $\Omega_M$ and $w$,
   the luminosities $\alpha$ and intrinsic dispersions $\sigma$ of SNe~Ia and non-Ia's,
   and the relative rate parameters $\theta_{r1}$ and $\theta_{r2}$.  Latent parameters
   include the type $T$ and redshifts of the host and neighboring galaxies $z$ and $z_N$
   for each transient.  The  parameters for distance modulus $\mu$ and
   luminosity $L$ are fixed by the other parameters.  The observables include the measured
   flux $\mathit{ADU}$, spectroscopic type and redshift $T_S$ and $z_S$, and
   the inferred host and neighbor redshifts $\theta_{G1}$ and $\theta_{G2}$.
   \label{toypgm:fig}}
\end{figure}

There are latent model parameters for each transient used internally
by the model:
\begin{itemize}
\item $z$: the cosmological redshifts.
\item $T$: the type, SN~Ia or non-Ia.
\end{itemize}

The transients are divided into samples defined by available data: {\it obs} are transients observed spectroscopically while active, from which $T_S=1$ are those typed
as SN~Ia and $T_S=0$ as non-Ia.  {\it mis} transients are not observed
spectroscopically when active.  The  {\it obs} and {\it mis} samples
are assumed to be drawn from the same underlying distribution, so that 
same SN~Ia and non-Ia parameters apply for both sets.

\subsubsection{Model for the {\it obs} set data}
The modeling of data from the {\it obs} set, which has
spectroscopic observations $T_S$ and redshift
$z_S$, is described as follows.  We are interested in
\begin{align}
p(\mathit{ADU}, {{T}}_S,{{z}}_S|  \Omega_M, w, \theta_r,\alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}})  &=\sum_T \int_z dz\,
p(\mathit{ADU}, {{T}}_S,{{z}}_S, T, z| \ldots)\\
&= \sum_T \int_z dz\,
p(\mathit{ADU}, {{T}}_S,{{z}}_S| T, z,\dots) p(T,z | \ldots)
\end{align}

A direct association of spectroscopic redshift and type are made so that
\begin{align}
p(z_S|z) &= \delta(z_S-z)\\
p(T_S|T) &= \delta(T_S-T),
\label{specz:eqn}
\end{align}
so that
\begin{align}
p(\mathit{ADU}, {{T}}_S,{{z}}_S|  \ldots) &= 
p(\mathit{ADU}| T=T_S, z=z_s,\dots) P(T_S| z= z_S , \ldots) p(z_S|\ldots).
\label{obs:eqn}
\end{align}
We remark that in a flux limited survey, the pdf's for the data are truncated versions of the universal
pdf's of the observables.

The first term describes the expected counts. Given the redshift
and transient type, the brightness of the underlying population
is modeled as being drawn from a standard candle
with luminosity $4\pi\alpha_X$ and intrinsic magnitude dispersion  $\sigma_X$:
\begin{equation}
\mathit{ADU}| T_S, z_S, \Omega_M, w, \theta_r, \alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}} \sim \mathcal{N}_{\ln}\left(\ln{\left(\frac{\alpha_{T_S}}{d_L^2(z_S;\Omega_M, w)}\right)}, \frac{\ln{10}}{2.5}\sigma_{T_S}\right),
\end{equation}
where $\mathcal{N}_{\ln}$ is the log-normal distribution and $d_L$ is the luminosity distance.

The subset of objects that enters our analysis comes from a truncated distribution:
\begin{itemize}
\item For transients in set $T_S=1$:
\begin{equation}
\mathit{ADU} | T_S=1, z_S, \Omega_M, w, \alpha_{Ia},\sigma_{Ia} \sim
\frac{\mathcal{N}_{\ln}\left(\ln{\left(\frac{\alpha_{Ia}}{d_L^2(z_S;\Omega_M, w)}\right)}, \frac{\ln{10}}{2.5}\sigma_{Ia}\right)}{P_{Ia}(\mathit{ADU} > \mathit{ADU}_0)}.
\label{adusnIa:eqn}
\end{equation}
\item For transients in set $T_S=0$:
\begin{equation}
\mathit{ADU} | T_S=0, z_S, \Omega_M, w, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}}\sim 
\frac{\mathcal{N}_{\ln}\left(\ln{\left(\frac{\alpha_{\mathit{non-Ia}}}{d_L^2(z_S;\Omega_M, w)}\right)}, \frac{\ln{10}}{2.5}\sigma_{\mathit{non-Ia}}\right)}{P_{non-Ia}(\mathit{ADU} > \mathit{ADU}_0)},
\label{adunonIa:eqn}
\end{equation}
\end{itemize}
where the normalization factor is the complementary cumulative distribution function
for the corresponding log-normal distribution.
It is understood that $p(\mathit{ADU}<\mathit{ADU}_0)=0$ is never evaluated.

The second term gives the type probability. In our model,
the underlying relative rates between SNe~Ia and non-Ia is linear in redshift,
 parameterized by the
its value at $z=0$ and $1.1 \times z_{max}$, 
\begin{equation}
\hat{\theta}_{r}=\theta_{r0}+z\left(\frac{\theta_{r1}-\theta_{r0}}{1.1 \times z_{max}}\right).
\end{equation}
The relative rate for SN~Ia discovery for the truncated distribution is
\begin{equation}
\theta_{r}=\frac{\hat{\theta}_{r}P_{Ia}(\mathit{ADU} > \mathit{ADU}_0)}{\hat{\theta}_{r}P_{Ia}(\mathit{ADU} > \mathit{ADU}_0) + (1-\hat{\theta}_{r})P_{non-Ia}(\mathit{ADU} > \mathit{ADU}_0)}.
\end{equation}
The spectroscopic classifications are drawn from
\begin{equation}
T_S | \Omega_M, w, \theta_{r0}, \theta_{r1} , \alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}} \sim \text{Bernoulli}(\theta_r).
\end{equation}
Although the underlying type distribution depends only on $ \theta_{r0}$ and $ \theta_{r1}$,  
the observed truncated distribution does depend on the other parameters through 
$P_{X}(\mathit{ADU} > \mathit{ADU}_0)$.

The third term is the redshift distribution of transient discoveries.
In our model we adopt
a redshift distribution of transients in the universe
\begin{equation}
P(z) \propto z^2.
\end{equation}
The redshift distribution of discovered transients is then
\begin{equation}
P(z) = \left(\int_{z_{min}}^{z_{max}} dz'\, z'^2\left( 
\hat{\theta}_{r}P_{Ia}(\mathit{ADU} > \mathit{ADU}_0) + (1-\hat{\theta}_{r})P_{non-Ia}(\mathit{ADU} > \mathit{ADU}_0)
\right)\right)^{-1}z^2.
\end{equation}

In the presented model, the data from different transients are independent.

\subsubsection{Model for the {\it mis} set data}

We now consider the modeling of data from the  {\it mis}
set that has no spectroscopic confirmation.
The redshift measurement is treated as follows:
For each transient there are two galaxies identified as
the potential host, one ($\theta_{G1}$) with high probability $P_{host}$
of being the host and the other
($\theta_{G2}$) at lower $(1-P_{host})$ probability.
In the model, the
true host redshift is $z$ and that of the interloping galaxy is $z_N$.
\begin{equation}
P(\theta_{G1},\theta_{G2}|z, z_N) =
	P_{host}\delta(z-\theta_{G1})\delta(z_N-\theta_{G2}) +
	(1-P_{host}) \delta(z-\theta_{G2})\delta(z_N-\theta_{G1}).
\end{equation}
As for the {\it obs} set, the redshifts of the supernova
 and neighbor are drawn from distributions $P(z) \propto z^2$, $P(z_N) \propto z^2_N$.

The data likelihood is 
\begin{multline}
p(\mathit{ADU}, \theta_{G1}, \theta_{G2} | \Omega_M, w, \theta_r, \alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}})  \\
= \sum_{T}\int_{z,z_N} dz\,dz_N \, p(\mathit{ADU, \theta_{G1}, \theta_{G2}}| z , z_N, T, \ldots)P(T| z, z_N,\ldots) p(z, z_N|\ldots)\\
 =\sum_{T} \int_{z,z_N} dz\,dz_N\, p(\mathit{ADU}| z ,T, \ldots)p(\theta_{G1}, \theta_{G2}| z , z_N) P(T| z, \ldots) p(z, z_N|\ldots)\\
= \sum_{T} P_{host} p(\mathit{ADU}| z=\theta_{G1} ,T, \ldots) P(T| z=\theta_{G1}, \ldots) p(\theta_{G1},\theta_{G2}|\ldots) \\
 +  (1-P_{host}) p(\mathit{ADU}| z=\theta_{G2} ,T, \ldots) P(T| z=\theta_{G2}, \ldots) p(\theta_{G2},
 \theta_{G1}|\ldots)\\
\propto \sum_{T} P_{host} p(\mathit{ADU}| z=\theta_{G1} ,T, \ldots) P(T| z=\theta_{G1}, \ldots) \\
 +  (1-P_{host}) p(\mathit{ADU}| z=\theta_{G2} ,T, \ldots) P(T| z=\theta_{G2}, \ldots),
\label{adumis:eqn}
\end{multline}
where we take advantage of the fact that
$p(\theta_{G1}, \theta_{G2} | \ldots) = p(\theta_{G2}, \theta_{G1} | \ldots)$ does
not have parameter dependence.

The distribution of $\mathit{ADU}$ for our sample is a truncated version of the
underlying distribution.  
For our analysis, we use the distribution in  Eqn.~\ref{adumis:eqn} normalized by
the complementary cumulative distribution function
\begin{equation}
P(\mathit{ADU}>\mathit{ADU}_0, \theta_{G1}, \theta_{G2}|\ldots) = \int_{\mathit{ADU}_0}^{\infty} p(\mathit{ADU}, \theta_{G1}, \theta_{G2}|\Omega_M, w, \theta_r, \alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}})d\mathit{ADU}
\end{equation}
for a detection threshold of $\mathit{ADU}_0$.

@@

\subsection{Analysis of Simulated Data}
The simulation of data from the model presented in \S\ref{example:sec}
and their analysis is implemented in Python and are available
on GitHub\footnote{{git@github.com:AlexGKim/abc.git}}.

The results that follow are based on the parameter
choices:
\begin{itemize}
\item Cosmological parameters $\Omega_M=0.28$, $w=-1$.
\item Type~Ia supernovae with intrinsic dispersion $\sigma_{SNIa}=0.1$.
\item Non-Ia supernovae 2 magnitudes fainter than SNe~Ia with intrinsic
dispersion $\sigma_{non-Ia}=1$.
\item $\theta_r=r_0 + (r_1-r_0)z/1.54$ for a constant relative fraction of SNe~Ia where $r_1=0.2$,
$r_0=0.95$.
\item 2000 transients distributed by co-moving volume in the range $0.1<z<1.4$.
\item Detection threshold at approximately the mean brightness SN~Ia at $z=1$.
\item $z_{min}=0.1/1.1$ and $z_{max}=1.4\times 1.1$ is the range of possible misassigned
host redshift.
\item $P_{host}=0.98$ for the probability of correct host-galaxy assignment.
\item Accumulated spectra for 0\%, 20\%, 60\%, and 100\% of discovered transients.
\end{itemize}
This setup is referred to as Model 1.

The Hubble diagram for a realization drawn from this model is shown in Figure~\ref{hd:fig}.
Type~Ia's are shown in blue and non-Ia's in red, those that fall below detection threshold
are faded.  The solid points
represent those objects spectroscopically typed for the case of 20\% follow-up.
The green X's represent the 2\% among those without spectroscopy that
are assigned an incorrect redshift.

>>>>

\documentclass[preprint,3p]{elsarticle}
\usepackage{aas_macros}
\usepackage{amsmath,amssymb}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{hyperref}

@@

Photometric supernova analysis has been proposed to  take advantage of the statistical weight imported by large numbers of supernova light curves, despite the lack of spectroscopy.
An example of such a program would proceed as follows: A rolling wide-field imaging
survey generates light curves for all transients in the survey area.  In near real-time,
a subset 
of transients from those images are selected for triggered spectroscopic observations
to obtain classifications and redshifts.  At the end of the survey, the
analysis sample is the subset of likely SN~Ia
candidates based on photometric data;  the analysis sample and its subset
with triggered spectroscopy are constructed so as to be drawn from the same parent distribution.
Spectroscopic redshifts are obtained for likely host galaxies.  The Hubble diagram of the
analysis sample is analyzed with the understanding that it may contain multiple populations.

The Sloan Digital Sky Survey-II Supernova Survey is an example of a program
that produced significant numbers of transients
with light curves but without spectroscopic typing \citep{2014arXiv1401.3317S}:
photometric SN analysis has been developed applied to these data \citep{2012ApJ...752...79H,
2013ApJ...763...88C}.   Such analyses are anticipated in the projected
supernova programs of 
DES \citep{2012ApJ...753..152B} and LSST \citep{2012arXiv1211.0310L}.

Current models used for supernova cosmology analysis account for unobserved quantities.
SNe~Ia are not standard but are ``standardizable'' candles, implying that the class may be composed of distinct subpopulations each with its own redshift-dependent rates, magnitude distributions, and secondary
observable (e.g.\ light-curve shape, color) distributions.  Bayesian hierarchical modeling is 
used to account for this extra layer of complexity  to determine characteristics of the SN~Ia population simultaneously
with the cosmology \citep{2011MNRAS.418.2308M,
2015arXiv150701602R}.  Consideration of missing classification is addressed
by
BEAMS \citep{2007PhRvD..75j3508K}, which  includes a non-Ia population in sample, whose luminosity distribution
is inferred along with the cosmology:
BEAMS was used in the analysis of SDSS-II SN data \citep{2012ApJ...752...79H}.

This article addresses the affect of missing spectroscopic typing in
a transient standard candle analysis, with the eventual goal of quantifying
requirements for the spectroscopic subsample in a broader photometric analysis.
The lack of spectroscopy leads to ambiguity in both the transient class and
in the host galaxy due to projected galaxies that appear close to the transient or
the relative low-surface brightness of the host.  The subsample with
spectroscopic typing informs the interpretation of objects lacking spectroscopy.
Differences between observed and
intrinsic distributions in a flux-limited sample can have important effects on the Hubble
diagram  (e.g.\ Malmquist bias) so are accounted for explicitly.
Though motivated by SNe~Ia, this article makes the simplifying assumption that they
are standard candles with random dispersion; the motivation of this article is to understand
the general effects of spectroscopic completeness in a manner that is not specific to
the choice of model for the underlying supernova heterogeneity.

\section{Data Model}
\label{model:sec}
The data in constructing a Hubble diagram and the fundamental parameters of the
model used to describe those data in this article are succinctly summarized in the notation
for the likelihood 
\begin{equation}
p(f, {{T}}_S,{{z}}_S, \theta_G|  \Omega_M, w, \theta_r,\alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}}),
\label{likelihood:eqn}
\end{equation}
and are sketched as the probabilistic graphical model in Figure~\ref{toypgm:fig}.
The measurements are
\begin{itemize}
\item $f$: the flux at peak brightness, taken 
to be bolometric to facilitate the treatment of
standard candles at different redshift;
\item ${{T}}_S$, ${{z}}_S$: the spectroscopic type and redshift available for a subset of transients;
\item $\theta_{G1}$, $\theta_{G2}$: the spectroscopic redshifts of the galaxies identified
as host and neighbor from imaging.
\end{itemize}
The model parameters are
\begin{itemize}
\item $\Omega_M$, $w$: the mass density and dark-energy equation of state parameters of a flat,
constant-$w$ dark-energy cosmology;
\item $\theta_{r1}$, $\theta_{r2}$:  the parameters
that describe the redshift-dependent relative fraction of SN~Ia's and non-Ia's;
\item $\alpha_{Ia}$, $\sigma_{Ia}$, $\alpha_{\mathit{non-Ia}}$, $\sigma_{\mathit{non-Ia}}$:
the luminosity (times $4\pi$) and intrinsic dispersion for SNe~Ia and non-Ia at peak brightness.
\end{itemize}
Not listed explicitly in Eqn.~\ref{likelihood:eqn} but shown in Figure~\ref{toypgm:fig} are latent model parameters for each transient:
\begin{itemize}
\item $z$: the cosmological redshifts of the host and neighboring galaxies;
\item $T$: the class of the objects, SN~Ia or non-Ia.
\end{itemize}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=4.5in]{/Users/akim/project/abc/results/toy_pgm.pdf} 
   \caption{Probabilistic Graphical Model for the  SN~Ia standard candle analysis. The
   parameters include the cosmological parameters $\Omega_M$ and $w$,
   the luminosities $\alpha$ and intrinsic dispersions $\sigma$ of SNe~Ia and non-Ia's,
   and the relative fraction parameters $\theta_{r1}$ and $\theta_{r2}$.  Latent parameters
   include the type $T$ and redshifts of the host and neighboring galaxies $z$ and $z_N$
   for each transient.  The  distance modulus $\mu$ and
   luminosity $L$ are fixed by the other parameters.  The observables include the measured
   flux $f$, spectroscopic type and redshift $T_S$ and $z_S$, and
   the inferred host and neighbor redshifts $\theta_{G1}$ and $\theta_{G2}$.
   The host and neighbor galaxies of all transients are inferred based a common galaxy catalog, 
   and so strictly are not independent from supernova to supernova.
   \label{toypgm:fig}}
\end{figure}

The analysis sample consists of transients that are discovered and pass some sample-selection criteria, for example a photometric-classification score.  While the studies in this
article  explicitly account for
the discovery criterion, they do not do so for the subsequent sample selection.
In other words, sample selection is assumed not to alter the shape of any of the pdf's.

The data for each transient
is modeled according to the following subsets:
{\it obs} are transients observed spectroscopically while active, from which $T_S=1$ are those typed
as SN~Ia and $T_S=0$ as non-Ia.  {\it mis} transients are not observed
spectroscopically when active.  The  {\it obs} and {\it mis} samples
are assumed to be drawn from the same underlying distribution, so that 
same SN~Ia and non-Ia parameters apply for both sets.

\subsection{Model for the {\it obs} set data}
\label{obs:sec}
The likelihood of data from the {\it obs} set, which has
spectroscopic observations $T_S$ and redshift
$z_S$, is described as follow.  The likelihood is
\begin{align}
p(f, {{T}}_S,{{z}}_S|  \Omega_M, w, \theta_r,\alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}})  &=\sum_T \int_z dz\,
p(f, {{T}}_S,{{z}}_S, T, z| \ldots)\\
&= \sum_T \int_z dz\,
p(f, {{T}}_S,{{z}}_S| T, z,\dots) p(T,z | \ldots).
\end{align}

A direct association of spectroscopic redshift and type are made
\begin{align}
p(z_S|z) &= \delta(z_S-z)\\
p(T_S|T) &= \delta(T_S-T),
\label{specz:eqn}
\end{align}
so that
\begin{align}
p(f, {{T}}_S,{{z}}_S|  \ldots) &= 
p(f| T=T_S, z=z_s,\dots) P(T_S| z= z_S , \ldots) p(z_S|\ldots).
\label{obs:eqn}
\end{align}

The first term describes the expected counts. Given the redshift
and transient type, the brightness of the underlying population
is modeled as being drawn from a standard candle
with mean luminosity $4\pi\alpha_X$ and a normal distribution in magnitude
space with intrinsic dispersion  $\sigma_X$:
\begin{equation}
f| T_S, z_S, \Omega_M, w, \theta_r, \alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}} \sim \mathcal{N}_{\ln}\left(\ln{\left(\frac{\alpha_{T_S}}{d_L^2(z_S;\Omega_M, w)}\right)}, \frac{\ln{10}}{2.5}\sigma_{T_S}\right),
\label{flux:eqn}
\end{equation}
where $\mathcal{N}_{\ln}$ is the log-normal distribution and $d_L$ is the luminosity distance.
Measurement uncertainty is taken to be negligible.
In a flux limited survey, the pdf's for the data are truncated versions of the underlying
pdf's.
\begin{itemize}
\item For SNe~Ia  $T_S=1$:
\begin{equation}
f | T_S=1, z_S, \Omega_M, w, \alpha_{Ia},\sigma_{Ia} \sim
\frac{\mathcal{N}_{\ln}\left(\ln{\left(\frac{\alpha_{Ia}}{d_L^2(z_S;\Omega_M, w)}\right)}, \frac{\ln{10}}{2.5}\sigma_{Ia}\right)}{P_{Ia}(f > f_0)}.
\label{adusnIa:eqn}
\end{equation}
\item For non-Ia's in set $T_S=0$:
\begin{equation}
f | T_S=0, z_S, \Omega_M, w, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}}\sim 
\frac{\mathcal{N}_{\ln}\left(\ln{\left(\frac{\alpha_{\mathit{non-Ia}}}{d_L^2(z_S;\Omega_M, w)}\right)}, \frac{\ln{10}}{2.5}\sigma_{\mathit{non-Ia}}\right)}{P_{non-Ia}(f > f_0)},
\label{adunonIa:eqn}
\end{equation}
\end{itemize}
where the normalization factor is the complementary cumulative distribution function
for the corresponding log-normal distribution.
It is understood implicitly that as all objects must pass the detection threshold,
the case of $f<f_0$ is never confronted in the analysis.

The second term gives the type probability. In our model,
the underlying relative rates between SNe~Ia and non-Ia is linear in redshift,
 parameterized by the
its values at $z=0$ and $1.1 \times z_{max}$, 
\begin{equation}
\hat{\theta}_r=\theta_{r0}+z\left(\frac{\theta_{r1}-\theta_{r0}}{1.1 \times z_{max}}\right).
\label{rate:eqn}
\end{equation}
The relative rate for SN~Ia discovery for the discovered (truncated) distribution is
\begin{equation}
\theta_r=\frac{\hat{\theta}_rP_{Ia}(f > f_0)}{\hat{\theta}_{r}P_{Ia}(f > f_0) + (1-\hat{\theta}_{r})P_{non-Ia}(f > f_0)}.
\end{equation}
The spectroscopic classifications are drawn from
\begin{equation}
T_S | \Omega_M, w, \theta_{r0}, \theta_{r1} , \alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}} \sim \text{Bernoulli}(\theta_r).
\end{equation}
Although the underlying type distribution depends only on $ \theta_{r0}$ and $ \theta_{r1}$,  
the observed distribution does depend on the other parameters through $\theta_r$'s dependence on
the two normalization factors
$P_{X}(f > f_0)$.

The third term is the redshift distribution of transient discoveries.
The model takes an underlying redshift distribution of transients in the universe
\begin{equation}
p(z) \propto z^2.
\end{equation}
The redshift distribution of discovered transients is then
\begin{equation}
p(z) = \frac{z^2\left(\hat{\theta}_{r}P_{Ia}(f > f_0) + (1-\hat{\theta}_{r})P_{non-Ia}(f > f_0)
\right)}{\int_{z_{min}}^{z_{max}} dz'\, z'^2\left( 
\hat{\theta}_{r}P_{Ia}(f > f_0) + (1-\hat{\theta}_{r})P_{non-Ia}(f > f_0)
\right)}.
\end{equation}

\subsection{Model for the {\it mis} set data}

Transients in the {\it mis}
set have no spectroscopic confirmation, and their data are treated as follows:
With no redshift from the transient itself, 
the redshift measurement is inferred from potential host galaxies.
In the analysis of this article each transient is associated with two galaxies, one 
with redshift $\theta_{G1}$
that has high probability $\pi_{host}$
of being the host and the other
with redshift $\theta_{G2}$ that has probability $(1-\pi_{host})$.
In the model, the
true host redshift is $z$ and that of the interloping galaxy is $z_N$.
\begin{equation}
P(\theta_{G1},\theta_{G2}|z, z_N) =
	\pi_{host}\delta(z-\theta_{G1})\delta(z_N-\theta_{G2}) +
	(1-\pi_{host}) \delta(z-\theta_{G2})\delta(z_N-\theta_{G1}).
\end{equation}
As for the {\it obs} set, the redshifts of the supernova
 and neighbor are drawn from underlying distributions $p(z) \propto z^2$, $p(z_N) \propto z^2_N$.

The underlying distribution is 
\begin{multline}
p(f, \theta_{G1}, \theta_{G2} | \Omega_M, w, \theta_r, \alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}})  \\
= \sum_{T}\int_{z,z_N} dz\,dz_N \, p(\mathit{ADU, \theta_{G1}, \theta_{G2}}| z , z_N, T, \ldots)P(T| z, z_N,\ldots) p(z, z_N|\ldots)\\
 =\sum_{T} \int_{z,z_N} dz\,dz_N\, p(f| z ,T, \ldots)p(\theta_{G1}, \theta_{G2}| z , z_N) P(T| z, \ldots) p(z, z_N|\ldots)\\
= \sum_{T} \pi_{host} p(f| z=\theta_{G1} ,T, \ldots) P(T| z=\theta_{G1}, \ldots) p(\theta_{G1},\theta_{G2}|\ldots) \\
 +  (1-\pi_{host}) p(f| z=\theta_{G2} ,T, \ldots) P(T| z=\theta_{G2}, \ldots) p(\theta_{G2},
 \theta_{G1}|\ldots)\\
\propto \sum_{T} \pi_{host} p(f| z=\theta_{G1} ,T, \ldots) P(T| z=\theta_{G1}, \ldots) \\
 +  (1-\pi_{host}) p(f| z=\theta_{G2} ,T, \ldots) P(T| z=\theta_{G2}, \ldots),
\label{adumis:eqn}
\end{multline}
where we take advantage of the fact that
$p(\theta_{G1}, \theta_{G2} | \ldots) = p(\theta_{G2}, \theta_{G1} | \ldots)$ has
no parameter dependence.  The probabilities  for the underlying flux and
relative rate distributions
 Eqns.~\ref{flux:eqn} and \ref{rate:eqn} introduced in \S\ref{obs:sec} apply.

The data likelihood is a truncated version of the
underlying distribution.  
For our analysis, we use the distribution in  Eqn.~\ref{adumis:eqn} normalized by
the complementary cumulative distribution function
\begin{equation}
p(f>f_0, \theta_{G1}, \theta_{G2}|\ldots) = \int_{f_0}^{\infty} p(f, \theta_{G1}, \theta_{G2}|\Omega_M, w, \theta_r, \alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}})df
\end{equation}
for a detection threshold of $f_0$.

@@

\section{Instantiation and Analysis of Simulated Data}
\subsection{Simulated Data}
Simulated data sets are realized and analyzed with the model of
 \S\ref{model:sec}.  The generative model of the first data set (Set 1) is
 identical to the model used for analysis, with the following parameter choices:
\begin{itemize}
\item Cosmological parameters $\Omega_M=0.28$, $w=-1$.
\item Type~Ia supernovae with intrinsic dispersion $\sigma_{SNIa}=0.1$.
\item Non-Ia supernovae 2 magnitudes fainter than SNe~Ia with intrinsic
dispersion $\sigma_{non-Ia}=1$.
\item $z_{max}=1.4$.
\item $\theta_{r0}=0.95$, $\theta_{r1}=0.2$ for the relative fraction of SNe~Ia
at $z=0$ and $z=1.1 \times z_{max}$ respectively.
\item 2000 transients total, including discovered and non-discovered,
in the range $0.1<z<z_{max}$.
\item Detection threshold at approximately the mean brightness SN~Ia at $z=1$.
\item Limits of neighbor-galaxy redshifts of $0.1/1.1$ and $1.1\times z_{max}$.
\item $\pi_{host}=0.98$ for the probability of correct host-galaxy assignment.
\item Accumulated spectra for 0\%, 20\%, 60\%, and 100\% of discovered transients.
\end{itemize}

The Hubble diagram for a realization drawn from this model is shown in Figure~\ref{hd:fig}
and a histogram of redshifts for discovered SNe~Ia and non-Ia in Figure~\ref{hist:fig}.
SNe~Ia's are shown in blue and non-Ia's in red, those that fall below detection threshold
are transparent.  The solid points
represent those objects spectroscopically typed for the case of 20\% follow-up.
The green X's represent the 2\% among those without spectroscopy that
are assigned an incorrect redshift.
